---

title : (Big) Data Analytics for Business
subtitle : Segmentation in the Boating Industry
author : T. Evgeniou and J. Niessing
job : INSEAD
widgets : []
mode : standalone 

---

## Every factor analysis has three basic decision points




Decide the number of factors

Choose an extraction method

Choose a rotation method

---

## Correlation matrix of the 29 attitudes as a pre-step


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE, results='asis'}
thecor=cor(ProjectDataFactor)
colnames(thecor)<-colnames(thecor)
rownames(thecor)<-rownames(thecor)
## printing the result in a clean-slate table
cat(renderHeatmapX(thecor, border=1))

```
** = correlation is significant at 1% level; * = correlation is significant at 5% level


---

## Number of Factors


There are three commen ways to select the number of factors:

1. All factors with eigenvalue larger than 1

2. Based on the cumulative variance we would like to explain (e.g. at least 50% for example).

3. Pick the number of factrors corresponding to the "elbow" of the scree plot



The first two approaches to select the number of factors/derived variables are based on statistical criteria and can be seen in the following table: 

---

## Eigenvalues and explained variance

```{r echo=FALSE, comment=NA, warning=FALSE, error=FALSE,message=FALSE,results='asis'}
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table
## print result in a clean-slate table
#Variance_Explained_Table[abs(Variance_Explained_Table)<MIN_VALUE]<-NA
print(xtable(Variance_Explained_Table, caption="Eigenvalues and Variance Explained", digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)

row=1:nrow(Variance_Explained_Table)
name<-paste("Component No:",row,sep="")
Variance_Explained_Table<-cbind(name,Variance_Explained_Table)
Variance_Explained_Table<-as.data.frame(Variance_Explained_Table)
colnames(Variance_Explained_Table)<-c("Components", "Eigenvalue", "Percentage_of_explained_variance", "Cumulative_percentage_of_explained_variance")

#print(gvisTable(Variance_Explained_Table, options=list(width=1200, height=400,allowHTML=TRUE,page='disable',pageSize=15,showRowNumber=FALSE,cssClassNames={tableCell='td.tableRow'})),formats=list(Eigenvalue="#,##") ,'chart')

m<-gvisTable(Variance_Explained_Table,options=list(width=1200, height=400,allowHTML=TRUE,page='disable'),formats=list(Eigenvalue="#.##",Percentage_of_explained_variance="#.##",Cumulative_percentage_of_explained_variance="#.##"))
print(m,'chart')
```


---

## A common basic visualization when running the factor analysis is the Screeplot

The screeplot depicts the eigenvalues of the factors and thus allows gaining another quick overview on how many factors should be retained:

---

## Extraction Method

The extraction method will produce factor loadings for every item on every extracted factor. 

There at least five extraction methods. The evidence supports the use of one of these two:

- principal axis factoring
- maximum likelihood approaches (i.e. canonical factoring)

The extraction and rotation steps are usually specificed in the same command in R (next subsection shows such commands).


---

## Rotation Method

Once an initial solution is obtained, the loadings are rotated. Rotation is a way of maximizing high loadings and minimizing low loadings so that the simplest possible structure is achieved.

There are two types of rotation:

- no rotation

- varimax, which is one of the key rotation methods used


--- 

## Based on the critrion 'explained variance (65%)' we have selected 10 factors that could now be used for the segmentation


```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
UnRotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none",score=TRUE)
factors_selected_tmp=factors_selected
if (factor_selectionciterion == "eigenvalue")
  factors_selected_tmp = which(UnRotated_Results$values > 1)

UnRotated_Factors<-UnRotated_Results$loadings[,factors_selected_tmp]
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Component",1:ncol(UnRotated_Factors),sep=" ")

UnRotated_Factors[abs(UnRotated_Factors) < MIN_VALUE]<-NA

print(xtable(UnRotated_Factors, caption=paste(paste("All PCA factors selected available with No rotation","(values less than",MIN_VALUE),"are not shown)"), digits=3), type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE,include.colnames = TRUE)
```

--- 

## Instead of the derived constructs we could also use 'a' single statement from each factor as a proxy for the factor

Usually researchers pick the attitudes with the highest factor loading

```{r echo=FALSE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
large_loadings <-as.matrix(rownames(Rotated_Factors)[apply(Rotated_Factors,2,which.max)],ncol=1)
rownames(large_loadings)<-paste("Factor",1:nrow(large_loadings), sep=" ")
print(xtable(large_loadings ,paste("Raw attributes with highest loading per factor:",data_name,sep=" "), digits=1,),type="html",html.table.attributes = "class='table table-striped table-hover table-bordered'",caption.placement="top",comment = FALSE,include.rownames = TRUE)

```
